% Use this template to write your solutions to COS 423 problem sets

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath, amsfonts, amsthm, amssymb, algorithm, graphicx, mathtools, xfrac}
\usepackage[noend]{algpseudocode}
\usepackage{fancyhdr, lastpage}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{pgfplots}
\usepackage[vmargin=1.20in,hmargin=1.25in,centering,letterpaper]{geometry}
\setlength{\headsep}{.50in}
\setlength{\headheight}{15pt}

% Landau notation
\DeclareMathOperator{\BigOm}{\mathcal{O}}
\newcommand{\BigOh}[1]{\BigOm\left({#1}\right)}
\DeclareMathOperator{\BigTm}{\Theta}
\newcommand{\BigTheta}[1]{\BigTm\left({#1}\right)}
\DeclareMathOperator{\BigWm}{\Omega}
\newcommand{\BigOmega}[1]{\BigWm\left({#1}\right)}
\DeclareMathOperator{\LittleOm}{\mathrm{o}}
\newcommand{\LittleOh}[1]{\LittleOm\left({#1}\right)}
\DeclareMathOperator{\LittleWm}{\omega}
\newcommand{\LittleOmega}[1]{\LittleWm\left({#1}\right)}

% argmin and argmax
\newcommand{\argmin}{\operatornamewithlimits{argmin}}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}

\newcommand{\calP}{\mathcal{P}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Exp}{\mathbb{E}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\sign}{\mathrm{sign\ }}
\newcommand{\abs}{\mathrm{abs\ }}
\newcommand{\eps}{\varepsilon}
\newcommand{\zo}{\{0, 1\}}
\newcommand{\SAT}{\mathit{SAT}}
\renewcommand{\P}{\mathbf{P}}
\newcommand{\NP}{\mathbf{NP}}
\newcommand{\coNP}{\co{NP}}
\newcommand{\co}[1]{\mathbf{co#1}}
\renewcommand{\Pr}{\mathop{\mathrm{Pr}}}

% theorems, lemmas, invariants, etc.
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{invariant}[theorem]{Invariant}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{property}[theorem]{Property}
\newtheorem{proposition}[theorem]{Proposition}

% piecewise functions
\newenvironment{piecewise}{\left \{\begin{array}{l@{,\ }l}}
{\end{array}\right.}

% paired delimiters
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
\DeclarePairedDelimiter{\len}{|}{|}
\DeclarePairedDelimiter{\set}{\{}{\}}

\makeatletter
\@addtoreset{equation}{section}
\makeatother
\renewcommand{\theequation}{\arabic{section}.\arabic{equation}}

% algorithms
\algnewcommand\algorithmicinput{\textbf{INPUT:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\algorithmicoutput{\textbf{OUTPUT:}}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}


% Formating Macros

\pagestyle{fancy}
\lhead{\sc \hmwkClass\ $\; \;\cdot \; \;$ \hmwkSemester\ $\; \;\cdot \; \;$
Problem \hmwkAssignmentNum.\hmwkProblemNum}
%\chead{\sc Problem \hmwkAssignmentNum.\hmwkProblemNum}
%\chead{}
\rhead{\em \hmwkAuthorName\ $($\hmwkAuthorID$)$\/}
\cfoot{}
\lfoot{}
\rfoot{\sc Page\ \thepage\ of\ \protect\pageref{LastPage}}
\renewcommand\headrulewidth{0.4pt}
\renewcommand\footrulewidth{0.4pt}

\fancypagestyle{fancycollab}
{
    \lfoot{\textit{Collaborators: \hmwkCollaborators}}
}

\fancypagestyle{problemstatement}
{
    \rhead{}
    \lfoot{}
}

%%%%%% Begin document with header and title %%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%%%% Header Information %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Shouldn't need to change these
\newcommand{\hmwkClass}{COS 255}
\newcommand{\hmwkSemester}{Spring 2016}

%%% Your name, in standard First Last format
\newcommand{\hmwkAuthorName}{Lukas Leung}
%%% Your NetID
\newcommand{\hmwkAuthorID}{lleung}

%%% The problem set number (just the number)
\newcommand{\hmwkAssignmentNum}{4}

%%% The problem number (just the number)
\newcommand{\hmwkProblemNum}{0}

%%% A list of your collaborators' NetIDs, separated by ", ".
%%% You can use a new line ("\\") in the middle to prevent a long
%%% list from overflowing.
\newcommand{\hmwkCollaborators}{}
%%% Sets the collaborator list to appear on the first page
\thispagestyle{fancycollab}

%%%%%%% begin Solution %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%% start Bank Cards %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Book 5.3: Bank Cards}
\textbf{Background:} \\
~\indent Suppose you're consulting for a bank that's concerned about
fraud detection, and they come to you with the following problem.
They have a collenction of $n$ bank cards that they've confiscated,
suspecting them of being used in fraud. Each bank card is a small
plastic object, containing a magnetic stripe with some encrypted
data, and it corresponds to a unique account in the bank. Each account
can have many bank cards corresponding to it, and we'll say that
two bank cards are $equivalent$ if they correspond to the same
account. It's very difficult to read the account number off a bank
card directly, but the bank has a high-tech "equivalence tester"
that takes two bank cards and, after performing some computations,
determines whether they are equivalent. \\
\textbf{Question:} \\
~\indent Among the collection of $n$ cards, is there a set of more
than $n/2$ of them that are all equivalent to one another? Assume
that the only feasible operations you can do with the cards are to
pick two of them and plug them in to the equivalence tester. Show
how to decide the answer to their question with only $(n\cdot log(n))$
invocations of the equivalence tester.
%%%%%%% end Problem %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage

%%%%%%% Mathematical Formulation %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Mathematical Formulation}
Given an input of $n$ bank cards, determine if $\exists$ a subset $M$
where all cards $c_i \in M$ are identical and $n/2\ \textless\ |M|$.
s.t. $c_i = c_j$.

%%%%%%% Algorithm %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Solution}
Important Confusing Data Structures:
\begin{itemize}
    \item BankCard[n] \textbf{bankCards} : ~ Store each Bank Card in
    each slot. We note that we are not able to sort this or do anything
    but access and swap positions i.e. can only access.
\end{itemize}

We will begin by making the observation that if $\exists$ a subset $M$
where all cards $c_i \in M$ are identical and $n/2\ \textless\ m\ =\ |M|$,
then in one of the halves $bankCards[0..(n/2)]$ or $bankCards[(n/2+1)..n]$
$\exists$ at least $m/2$ identical cards. This algorithm will implement
this recursively in order to determine whether or not the afformentioned
statement holds. $Note: for\ the\ sake\ of\ this\ algorithm\ I\ will\ use\ BankCard[\ ]\
to\ pass\ through$ $but\ the\ actual\ algorithm\ would\ use\ a\ global\
array\ and\ pass\ the\ corresponding\ indices$

\begin{algorithm}[H]
\caption{BankCard}
\begin{algorithmic}
    \Procedure{Equivalence}{BankCard[] bankCards}
        \State $n \gets$ \Call{sizeOf}{bankCards}
        \State $m \gets$ n/2
        \If{n == 1}
            \State return bankCard[0]
        \Else{\ n == 2}
            \If{bankCards[0].\Call{equals}{bankCards[1]}}
                \State return bankCard[0]
            \EndIf
        \EndIf
        \State $bankCards1, bankCards2 \gets$ bankCards[0..m], bankCards[(m+1)..n]
        \State $card1, card2 \gets$ \Call{Equivalence}{bankCards1}, \Call{Equivalence}{bankCards2}
        \If{card1 is a card}
            \State test card1 against all other cards in bankCards and numberEqual++ when equal
            \If{numberEqual == GOAL\_SIZE}
                \State return card1
            \EndIf
        \EndIf
        \If{card2 is a card}
            \State test card2 against all other cards in bankCards and numberEqual++ when equal
            \If{numberEqual == GOAL\_SIZE}
                \State return card2
            \EndIf
        \EndIf
        \State return bankCards
    \EndProcedure
\end{algorithmic}
\end{algorithm}

If at the end of this method no card has been returned $\implies$
there is no matching of size n/2



%%%%%%% Correctness %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Correctness}
%%%%%%% PROPOSITION 1 %%%%%%%%%%%%%%%
\begin{proposition}
~ \\ \indent Propose that the algorithm will determine if $\exists$
a subset $M$ where all cards $c_i \in M$ are identical and
$n/2\ \textless\ m\ =\ |M|$.

\end{proposition}

\begin{proof}
~ \\ \indent This holds because if more than $n/2$ cards are equivalent,
one of the halves $bankCards[0..(n/2)]$ or $bankCards[(n/2+1)..n]$
$\exists$ more than $m/2$ identical cards. $\therefore$ one of the two
recursive calls must return a card equivalent to the whole set’s
majority equivalence $\therefore$ this algorithm compares all returned
cards to the whole set, so then the majority equivalence will be found.
\end{proof}


%%%%%%% Analysis %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Analysis}
For the following analysis, we will say that..

%%%%%%% PROPOSITION 1 %%%%%%%%%%%%%%%
\begin{proposition}
\label{numq}
The \underline{space complexity} of this algorithm is \textbf{O(N)}
\end{proposition}

\begin{proof}
~ \\ \indent This is due to the fact that all of our data is stored
in the array containing all the cards. Everything else is pointers
which are being initialized recursively so at most $3\cdot log(N)$
of them will be active at a time.
\begin{center}
    Giving us a space complexity of \textbf{O(N)}
\end{center}
\end{proof}

%%%%%%% PROPOSITION 2 %%%%%%%%%%%%%%%
\begin{proposition}
\label{numq}
The \underline{time complexity} of this algorithm is \textbf{O(N$\cdot$log(N))}
\end{proposition}

\begin{proof}
This is the case because our algorithm calls the $Equivalence$ method,
where N cards = T(N), as two recursive calls (each of size n/2) and
at most 2n tests for the two returned cards at each level. So T(n)
$\cong$ 2T(n/2) + 2n = O(n log n) from class.
\begin{center}
    Giving us a time complexity of \textbf{O(N$\cdot$log(N))}
\end{center}
\end{proof}

%%%%%%% end Bank Cards %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
%%%%%%% start Video Streams %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Book 4.12: Video Streams}
\textbf{Background} \\
~\indent Suppose you have $n$ video streams that need to be sent,
one after another, over a communication link. Stream $i$ consists
of a total of $b_i$ bits that need to be sent, at a constant rate
over a period of $t_i$ seconds. You cannot send two streams at the
same time, so you need to determine a $schedule$ for the streams:
an order in which to send them. Whichever order you choose, there
cannot be any delays between the end of one stream and the start
of the next. Suppose your schedule starts at time 0 (and therefore
ends at time $\sum_{i=1}^{n} t_i$, whichever order you choose). We
assume that all the values $b_i$ and $t_i$ are positive integers. \\
\indent Now, because you're just one user, the link does not want
you taking up too much bandwidth, so it imposes the following
constraint, using a fixed parameter $r$.

~ \\ \indent \textit{($\ast$) For each natural number t
$\textgreater$ 0, the total number of bits you send over the \\
~ \indent time interval from 0 to t cannot exceed r$\cdot$t.} \\

Note that this constraint is only imposed for time intervals that
start at 0, $not$ for time intervals that start at any other value.
We say that a schedule is $valid$ if it satisfies the constraint
($\ast$) imposed by the link. \\

\textbf{Questions:} \\
~\indent Given a set of $n$ streams, each specified by its number
of bits $b_i$ and its time duration $t_i$, as well as the link
parameter r, determine whether $\exists$ a valid schedule.

\subsubsection{Question 1:}
\begin{center}\textit{Claim: $\exists$ a valid schedule $\iff$
each stream i satisfies $b_i\leq r\cdot t_i$.}\end{center}
Decide whether you think the claim is true or false, and give a
proof of either the claim or its negation.

\subsubsection{Question 2:}
Give an algorithm that takes a set of $n$ streams, each specified
by its number of bits $b_i$ and its time duration $t_i$, as well as
the link parameter $r$, and determines whether there exists a valid
schedule. The running time of your algorithm should be polynomial
in $n$.

%%%%%%% end Problem %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
%%%%%%% first Problem %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Question 1:}
\textbf{False.} \\
Proof by example, giving the case:  $s_1$ = (1,1), $s_2$ = (2,1)
and $r$ = 1. We see here that this is clearly an acceptable order as
there is no space and for \underline{t=1} $\implies$ $b_1 \leq r\cdot t$
i.e. $1 \leq 1\cdot 1$ so $\ast$ holds. Similarly for \underline{t=2}
$\implies$ $b_2 \leq r\cdot t$ i.e. $2 \leq 1\cdot 2$ so $\ast$ holds.
However this will fail the case since for $s_2$, $b_2 \nleq r\cdot t_2$
i.e. $2 \nleq 1\cdot 1 \implies$ fails the case. $\square$

\subsection{Question 2:}
%%%%%%% Mathematical Formulation %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Mathematical Formulation}
Given an input of $n$ video streams, $s_1 = (b_1,t_1), s_2 = (b_2, t_1),
..., s_n = (b_n,t_n)$ where $b_i$ = the load of stream $i$ and $t_i$ =
the duration for how long it takes to transmit all data. We will say
that in the correct order, there cannot be any delays between the end
of one stream and the start of the next. i.e. time starts t=1 and goes
to $\sum_{i=1}^{n} t_i$. We will denote the current time as $T_i$
for when each step $i$ begins where $T_1 = 1$. Additionally there is a
rule that $b_i \leq r\cdot T_i$ for each stream $i$ where $r$ is a
fixed parameter.

%%%%%%% Algorithm %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Solution}
Important Confusing Data Structures:
\begin{itemize}
    \item VideoStream[n] \textbf{streams} : ~ Holds the video streams
    storing them as objects able to retrieve their $(d_i, t_i)$
\end{itemize}

\noindent The main functionality of this is to sort the video streams
by $d_i/t_i$, and then check to see if the condition is met. If at
any point it is not, then there does not exist a valid arrangement.

\begin{algorithm}[H]
\caption{Method}
\begin{algorithmic}
    \Procedure{sortStreams}{streams, r}
        \State \Call{sort}{streams, Comparator()}
        \State $T \gets$ 1
        \For{$s_i \in streams$}
            \If{$b_i\ \textgreater\ r\cdot T$}
                \State return null
            \EndIf
            \State $T += t_i$
        \EndFor
        \State return streams
    \EndProcedure
    \Procedure{compare}{$d_a,t_a,d_b,t_b$}
        \State $pos_a \gets d_a/t_a$
        \State $pos_b \gets d_b/t_b$
        \State return $pos_a - pos_b$
    \EndProcedure
\end{algorithmic}
\end{algorithm}


%%%%%%% Correctness %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Correctness}
%%%%%%% PROPOSITION 1 %%%%%%%%%%%%%%%
\begin{proposition}
~ \\ \indent Propose that by sorting via $d_i/t_i$ we gaurantee
$\exists$ a valid ordered input.
\end{proposition}

\begin{proof}
~ \\ \indent The motivation behind this is using the equality
$d_i \leq r\cdot T_i \implies d_i/T_i \leq r$ must also hold. Therefore,
we see that the $i^{th}$ stream is depenent on the ratio between the
load and the current time. $\therefore$ We develop our rational this
way by saying that if we order them then in the order of their
ratio, then those with lower $b_i$ compared to $t_i$ would appear
first, i.e. $d_1/t_1 \leq d_2/t_2 \leq ... \leq d_n/t_n$. If $\exists$
a valid output then this will give us a valid output. This is due
to the fact that if you were to rearrange any part, then the ratios
would be inverted and then streams would be worse off.
\end{proof}


%%%%%%% Analysis %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Analysis}
\begin{center}Let $N$ be the total number of streams being considered.\end{center}

%%%%%%% PROPOSITION 1 %%%%%%%%%%%%%%%
\begin{proposition}
\label{numq}
The \underline{space complexity} of this algorithm is \textbf{O(N)}
\end{proposition}

\begin{proof}
~ \\ \indent This is due to the fact that all of our data is stored
within the object array of size $N$ and two integers $T\ and\ r$.
\begin{center}
    Giving us a space complexity of \textbf{O(N)}
\end{center}
\end{proof}

%%%%%%% PROPOSITION 2 %%%%%%%%%%%%%%%
\begin{proposition}
\label{numq}
The \underline{time complexity} of this algorithm is \textbf{O(N$\cdot$log(N))}
\end{proposition}

\begin{proof}
This is the case because our algorithm takes ($N\cdot log(N)$) to sort
based off of this comparator and then linearly searches through, ($N$),
doing the single check statement (1)
\begin{center}
    Giving us a time complexity of \textbf{O(N$\cdot$log(N))}
\end{center}
\end{proof}

%%%%%%% end Video Streams %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



%%%%%%% end Solution %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}